Local Sensitivity Analysis of a Supercritical Extraction Model

### 2025-01-25 Article_Sensitivity_Analsysis-Chamomile.tex
25.01

	% ===================================================
	% ABSTRACT
	% ===================================================
	\begin{abstract}
		This study investigates the process of chamomile oil extraction from chamomile flowers. A parameter-distributed model, consisting of a set of partial differential equations, was used to describe the governing mass transfer phenomena between solid and fluid phases under supercritical conditions using carbon dioxide as the solvent. The concept of quasi-one-dimensional flow was applied to reduce the number of spatial dimensions. The flow of carbon dioxide is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. The physical properties of the solvent were estimated using the Peng-Robinson equation of state. Laboratory experiments were conducted under various, but constant operating conditions of $30 - 40~^\circ C$, $100 - 200$ bar and $3.33-6.67 \cdot 10^{-5}$ kg/s. The local sensitivity analysis method was applied to evaluate the robustness of the model parameters by investigating the impact of infinitesimally small changes in the model parameters and controls on the model outputs. This study focuses on analysing the effect of pressure on the model state space and extraction yield.

@@

	The applications of supercritical carbon dioxide are not limited only to an extraction process but can also be used for impregnation, as described by \citet{Weidner2018}, \citet{Machado2022} or \citet{Fathi2022}. Impregnation is defined as modifying the properties of bulk substances by physically or chemically binding/adsorbing impregnates to a bulk material or surface, such as the hydrophobization of surfaces. The main advantage of using supercritical CO$_2$ is that, after depressurization, it desorbs from the surface and evaporates, leaving a solvent-free product. On the other hand, the main disadvantage of using carbon dioxide for impregnation is the low solubility of many drugs of interest.

	Another application of supercritical CO$_2$ is nanoparticle formation, as investigated by \citet{Padrela2018}, \citet{Franco2021}, \citet{SaadatiArdestani2020} or \citet{Sodeifian2022}. Supercritical carbon-dioxide-assisted technologies enable the production of different morphologies of different sizes, including nanoparticles and nanocrystals, by modulating the operating conditions. Supercritical fluid-based processes have advantages over techniques conventionally employed to produce nanosized particles or crystals, such as reduced use of toxic solvents. Moreover, the CO$_2$ is removed from the final product by simple depressurization.

@@

@@

	%\subfile{Sections/Literature_Review}
	%This study aims to analyse the influence of changes in operating conditions on the process model described by \citet{Sliczniuk2024}. For that, a sensitivity analysis is applied to examine the impact of model parameters or controls on the model output. The results of this sensitivity analysis can be used to identify sources of uncertainty, simplify the model or detect errors by revealing unexpected relationships between inputs and outputs. Various sensitivity analysis methods are available, including:

@@

	The results of this sensitivity analysis are valuable for multiple purposes. First, they help identify sources of uncertainty within the model, highlighting which parameters most significantly influence predictions. Second, they provide insights for potential model simplification, allowing the exclusion of non-essential parameters without compromising accuracy. Finally, the analysis can detect inconsistencies or errors by uncovering unexpected relationships between inputs and outputs, which may signal flaws in the model's assumptions or structure.

	A range of sensitivity analysis methods can be utilized to achieve these objectives, including:}

@@

	\citet{Hatami2024} performed a one-factor-at-a-time sensitivity analysis to assess the response of net present value (NPV) to variations in technical and economic variables. Their study consisted of two parts. The first part examined how net present value is influenced by changes in individual technical and economic parameters, keeping the extractor volume constant at 300 L. The second part investigated the effects of varying the extractor volume (from 1 to 600 L) on the project's overall profitability. The authors found that the raw material price, discount rate and residence time had the biggest impact on NPV (\citet{Hatami2024}).

@@

	One disadvantage of using local sensitivity coefficients is that their reflect the absolute changes, which might be difficult to compare among other models or scenarios. The normalized local sensitivity coefficients, defined as $\tilde{S}(x;\Theta) = \frac{d G(x;\Theta)}{d\Theta} \frac{\Theta}{G(x;\Theta)}$, which can be interpreted as the percentage change in the output per percentage change in the parameter, making it a unitless. Although, the interpretation is straightforward it should be noted that extrapolation for large change of the parameter might leads to incorrect results. In the case of dynamic systems, the outcome of local sensitivity analysis quantifies how the state of the system or its output varies over time with respect to small perturbations in parameters. If the perturbation is too large, then the definition of a derivative (or sensitivity equations) does not hold unless the system is linear. Consequently, the conclusions can not be scaled up to any size of parameter perturbation. Such a property of the local sensitivity analysis, makes it difficult to recreate in a laboratory conditions due to finite precision of machines.

	{\color{blue}Figure \ref{fig:Sensitivty_P_y_norm} illustrates the time evolution of the normalized sensitivity coefficients of the extraction yield. Unlike the raw sensitivities shown in Figure \ref{fig:Sensitivty_P_y}, the normalized sensitivities do not start from zero. This is due to the absence of the exact initial time point on the logarithmic time axis. The pressure change impacts the system instantaneously and uniformly, leading to a rapid initial change in the normalized sensitivities. This behavior is analogous to a step change, where the system's response to the perturbation is immediate. Initially, the sensitivity curves remain flat. The increase in normalized sensitivity begins when the solute reaches the system's outlet and becomes detectable. After reaching their respective maxima, the sensitivity curves decrease as the concentration gradient driving the extraction diminishes over time. Ultimately, all curves approach zero, reflecting that all the solute was removed from the solid phase.}

@@

	Sensitivity analysis is a tool used to understand how the parameters of the model affect its output. The presented formulation involves derivative-based local sensitivity analysis of the model solution with respect to selected parameters and controls. This work implemented automatic differentiation to derive the sensitivity equations. Local sensitivity analysis techniques consider only a small region of parameter space, and the conclusions derived from such an analysis are limited to local conditions.

@@

	It was observed that the system responses at low pressures were stronger than those at higher pressures. This behaviour can be explained by the rapid changes in solvent properties occurring near the critical point of CO$_2$.

	\sout{
	Local sensitivity analysis results provide valuable information by identifying which parameters are influential and how these influences change over time. One application of sensitivity analysis is to spot discrepancies between model predictions and actual experimental results. Identifying unexpected system responses prompts further investigation and refinement of the process model.  The sensitivity analysis results can be used to design future experiments by indicating which parameters should be varied and monitored. Moreover, the parameters which are characterized by low sensitivity can be the subject of model refinement. }

@@

	\newpage

>>>>

	% ===================================================
	% ABSTRACT
	% ===================================================
	\begin{abstract}
		This study investigates the process of chamomile oil extraction from chamomile flowers. A parameter-distributed model, consisting of a set of partial differential equations, was used to describe the governing mass transfer phenomena between solid and fluid phases under supercritical conditions using carbon dioxide as the solvent. The concept of quasi-one-dimensional flow was applied to reduce the number of spatial dimensions. The flow of carbon dioxide is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. The physical properties of the solvent were estimated using the Peng-Robinson equation of state. Laboratory experiments were conducted under various, but constant operating conditions of $30 - 40~^\circ C$, $100 - 200$ bar and $3.33-6.67 \cdot 10^{-5}$ kg/s. The local sensitivity analysis method was applied to evaluate the robustness of the model parameters by investigating the impact of infinitesimally small changes in the model parameters and controls on the model outputs. This study focuses on analysing the effect of pressure on the model state space and extraction yield. {\color{blue}It was found that the model is the most sensitive to the controls if operated close to the critical point.}

@@

	The applications of supercritical carbon dioxide are not limited only to an extraction process but can also be used for impregnation, as described by \citet{Weidner2018} or \citet{Machado2022}. Impregnation is defined as modifying the properties of bulk substances by physically or chemically binding/adsorbing impregnates to a bulk material or surface, such as the hydrophobization of surfaces. The main advantage of using supercritical CO$_2$ is that, after depressurization, it desorbs from the surface and evaporates, leaving a solvent-free product. On the other hand, the main disadvantage of using carbon dioxide for impregnation is the low solubility of many drugs of interest.
	{\color{blue}The study of \citet{Ameri2020} investigates the loading of lansoprazole into polymers using supercritical carbon dioxide and examines how various parameters, such as temperature, pressure, and time, affect the drug loading efficiency. The results indicate that increasing any of these parameters enhances drug loading, with temperature having the most significant impact.	The study of \citet{Fathi2022} explores the use of supercritical carbon dioxide to enhance the bioavailability of ketoconazole by impregnating it into water-soluble polymers, specifically polyvinylpyrrolidone and hydroxypropyl methylcellulose. Utilizing a Box-Behnken design, the researchers optimized the impregnation process by varying pressure (150–250 bar), temperature (308–328 K), and time (60–180 minutes), achieving increase in drug loadings ranging.}

	Another application of supercritical CO$_2$ is nanoparticle formation, as investigated by \citet{Padrela2018}, \citet{Franco2021} or \citet{Sodeifian2022}. Supercritical carbon-dioxide-assisted technologies enable the production of different morphologies of different sizes, including nanoparticles and nanocrystals, by modulating the operating conditions. Supercritical fluid-based processes have advantages over techniques conventionally employed to produce nanosized particles or crystals, such as reduced use of toxic solvents. Moreover, the CO$_2$ is removed from the final product by simple depressurization. 
	{\color{blue} \citet{Sodeifian2018} investigates the solubility of Letrozole, a poorly water-soluble anticancer drug, in supercritical carbon dioxide with and without menthol as a solid co-solvent. The addition of menthol increased Letrozole's solubility by 7.1 times compared to supercritical CO$_2$ alone. Using the rapid expansion of supercritical solutions with solid co-solvent method, the average particle size of Letrozole was reduced to the nanoscale. Temperature was found to have the most significant impact on nanoparticle size reduction, while pressure had the least effect. 
	The study of \citet{SaadatiArdestani2020} explores the preparation of phthalocyanine green nano pigment using supercritical carbon dioxide as an antisolvent. The researchers employed the gas antisolvent technique to achieve nano-sized particles of the pigment. This method offers a potential approach for producing nano pigments with enhanced properties. The findings contribute to the development of advanced materials in the field of pigments and coatings.
	The study of \citet{Sodeifian2019} investigates the production of amiodarone hydrochloride nanoparticles using an ultrasonic-assisted rapid expansion of supercritical solution into a liquid solvent method. By optimizing parameters such as pressure, temperature, and polymeric stabilizer concentration, the researchers achieved significant particle size reduction. Characterization techniques confirmed the successful formation of nanoparticles with improved properties.
	}

@@

	{\color{blue}Depending on the needs one of two approaches can be considered while developing a mathematical model. A model, which is based on a multiple regression can be used if the relation between inputs and outputs can be simplified. \citet{Sodeifian2017a} investigated an influence of pressure, temperature and particle size on the extraction efficiency of oil from dracocephalum kotschyi boiss seed. The second order polynomial model was applied to obtain the corresponding response surface and to identify the optimum operating conditions. 
	The study of \citet{Sodeifian2017b} investigates the of essential oil from Eryngium billardieri, focusing on optimizing extraction conditions and developing a mathematical model based on the f the second order polynomial to predict the process yield. The researchers employed a simulated annealing algorithm to optimize parameters such as pressure, temperature, and extraction time, aiming to maximize oil efficiency.

	Alternatively, a more complex model can be derived and applied to cover not only the input-output relations, but the phenomena occurring in the system. This approach allows for more detailed representation the system behaviour, but requires better understanding of the mass transfer and more rigorous experiments.}

@@

@@

	The results of this sensitivity analysis are valuable for multiple purposes. First, they help identify sources of uncertainty within the model, highlighting which parameters most significantly influence predictions. Second, they provide insights for potential model simplification, allowing the exclusion of non-essential parameters without compromising accuracy. Finally, the analysis can detect inconsistencies or errors by uncovering unexpected relationships between inputs and outputs, which may signal flaws in the model's assumptions or structure. }

	A range of sensitivity analysis methods can be utilized to achieve these objectives, including:

@@

	\citet{Hatami2024} performed a one-factor-at-a-time sensitivity analysis to assess the response of net present value (NPV) to variations in technical and economic variables. Their study consisted of two parts. The first part examined how net present value is influenced by changes in individual technical and economic parameters, keeping the extractor volume constant at 300 L. The second part investigated the effects of varying the extractor volume (from 1 to 600 L) on the project's profitability. The authors found that the raw material price, discount rate and residence time had the biggest impact on NPV.

	{\color{blue}The paper of \citet{Maly1996} addresses the development of numerical methods and software tools for sensitivity analysis in differential-algebraic equation systems. The paper discusses the mathematical foundations of the local sensitivity analysis methods employed, providing a thorough analysis of their performance. The authors presented variety of problems and to prove the efficient of their method.

	In his \citet{Turanyi1990} presents a comprehensive review of sensitivity analysis methods applied to spatially homogeneous, constant-parameter chemical kinetic systems. The work delves into various sensitivity analysis techniques, including concentration sensitivity, rate sensitivity, and feature sensitivity analysis. The paper explores practical applications of these methods in chemical kinetics, demonstrating their utility in understanding and predicting system behavior.

	\citet{Fishtik1997} applied the local sensitivity analysis of an multiple chemical reaction system with several pure condensed phases and one ideal gas phase. It was shown that the system's response may be presented as a sum of contributions coming only from response equilibria defined by gas phase species. The effect of temperature, pressure and initial amounts of species is analyzed for gasification of coal.

	\citet{Saltelli2005} provide a comprehensive review of sensitivity analysis methods applicable to chemical modeling. The authors discuss both local and global techniques, emphasizing the importance of global sensitivity analysis in understanding complex chemical systems. The manuscript highlight the role of variance-based methods, in quantifying the influence of input parameters on model outputs.

	Unlike previous studies, this work focuses on derivative-based local sensitivity analysis to evaluate the robustness and predictive accuracy of a mathematical model for chamomile oil extraction under varying operational conditions. The analysis specifically examines the impact of operating condition changes, providing new insights into model development and optimizing extraction performance.}

@@

	{\color{blue}One disadvantage of using local sensitivity coefficients is that their reflect the absolute changes, which might be difficult to compare among other models or scenarios. The normalized local sensitivity coefficients, defined as $\tilde{S}(x;\Theta) = \frac{d G(x;\Theta)}{d\Theta} \frac{\Theta}{G(x;\Theta)}$, which can be interpreted as the percentage change in the output per percentage change in the parameter, making it a unitless. Although, the interpretation is straightforward it should be noted that extrapolation for large change of the parameter might leads to incorrect results. In the case of dynamic systems, the outcome of local sensitivity analysis quantifies how the state of the system or its output varies over time with respect to small perturbations in parameters. If the perturbation is too large, then the definition of a derivative (or sensitivity equations) does not hold unless the system is linear. Consequently, the conclusions can not be scaled up to any size of parameter perturbation. Such a property of the local sensitivity analysis, makes it difficult to recreate in a laboratory conditions due to finite precision of machines.

	Figure \ref{fig:Sensitivty_P_y_norm} illustrates the time evolution of the normalized sensitivity coefficients of the extraction yield. Unlike the raw sensitivities shown in Figure \ref{fig:Sensitivty_P_y}, the normalized sensitivities do not start from zero. This is due to the absence of the exact initial time point on the logarithmic time axis. The pressure change impacts the system instantaneously and uniformly, leading to a rapid initial change in the normalized sensitivities. This behavior is analogous to a step change, where the system's response to the perturbation is immediate. Initially, the sensitivity curves remain flat. The increase in normalized sensitivity begins when the solute reaches the system's outlet and becomes detectable. After reaching their respective maxima, the sensitivity curves decrease as the concentration gradient driving the extraction diminishes over time. Ultimately, all curves approach zero, reflecting that all the solute was removed from the solid phase.}

@@

	Sensitivity analysis is a tool used to understand how the parameters of the model affect its output. The presented formulation involves derivative-based local sensitivity analysis of the model solution with respect to selected parameters and controls. This work implemented automatic differentiation to derive the sensitivity equations. The results of the local sensitivity analysis and model specific and consider only a small region of parameter space, and the conclusions derived from such an analysis are limited to local conditions.

@@

	It can be concluded that the system responses at low pressures were stronger than those at higher pressures. This behaviour can be explained by the rapid changes in solvent properties occurring near the critical point of CO$_2$.

@@

	\clearpage
